PineAPPL: combining EW and QCD corrections\\ for fast evaluation of LHC processes

### 2020-08-27 paper/sec-conclusion.tex
Cosmetic changes on the final line of the conclusions
Polish introduction and conclusions

\section{Conclusions}
\label{sec:conclusion}

The systematic inclusion of EW corrections in accurate theoretical computations
for several LHC processes is becoming more and more important in order to
match the increasing precision of the data. In this paper we simplified the
computational aspect of this task, building upon the automation of QCD and EW
computations pioneered in recent years~\cite{Kallweit:2014xda,Biedermann:2017yoi,Frederix:2018nkq}.
Specifically we developed \textsc{PineAPPL}, a new library that stores perturbative calculations from an external Monte Carlo generator in a PDF-independent way using interpolation grids.
This offers the advantage of fast a posteriori convolutions with PDFs, for example to study the uncertainties coming from different PDF sets and/or the strong coupling constant $\alphas$, and to determine the PDFs themselves, a task for
which fast-interpolation grids are fundamental.
We tested \textsc{PineAPPL} together with \textsc{mg5\_aMC} and found a precision of \numrange{e-4}{e-5} relative to the MC result, which is excellent for all
foreseeable practical purposes.
Although we used \textsc{mg5\_aMC}, we note that \textsc{PineAPPL} is not tied in any way to a specific Monte Carlo generator, and can be easily interfaced with any of them.

@@

The \textsc{PineAPPL} library is available at \url{https://n3pdf.github.io/pineappl}.

>>>>

\section{Conclusions and Outlook}
\label{sec:conclusion}

The systematic inclusion of EW corrections in accurate theoretical computations
for several LHC processes is becoming more and more important in order to
match the increasing precision of the data. In this paper we simplified the
computational aspect of this task, building upon the automation of QCD and EW
computations pioneered in recent years~\cite{Kallweit:2014xda,Biedermann:2017yoi,Frederix:2018nkq}.
Specifically we developed \textsc{PineAPPL}, a new library that stores perturbative calculations from an external Monte Carlo generator in a PDF-independent way using interpolation grids.
This offers the advantage of fast a posteriori convolutions with PDFs, for example to study the uncertainties coming from different PDF sets and/or the strong coupling $\alphas$, and to determine the PDFs themselves, a task for
which fast-interpolation grids are fundamental.
We tested \textsc{PineAPPL} together with \textsc{mg5\_aMC} and found a precision of \numrange{e-4}{e-5} relative to the MC result, which is excellent for all
foreseeable practical purposes.
Although we used \textsc{mg5\_aMC}, we note that \textsc{PineAPPL} is not tied in any way to a specific Monte Carlo generator, and can be easily interfaced with any of them.

@@

\vspace{0.5cm}
\hrule
\begin{center}
The \textsc{PineAPPL} library is available at \url{https://n3pdf.github.io/pineappl}.
\end{center}

### 2020-08-27 paper/sec-introduction.tex
Polish introduction and conclusions

With the recent completion of Run II, the Large Hadron Collider (LHC) has 
accumulated data from an integrated luminosity of approximately 
\SI{150}{\per\femto\barn}~\cite{Mangano:2020icy}. This represents only a small fraction of
the anticipated \SI{3000}{\per\femto\barn} that will eventually be recorded in the
forthcoming twenty years of LHC operation. Nevertheless the statistical 
uncertainty of the data has already shrunk to unprecedentedly small values,
typically \SI{1}{\percent} or less, a fact that will allow for precision tests of the
Standard Model (SM) and for indirect searches of New Physics only if 
theoretical predictions become comparatively precise. This entails the 
computation of additional higher-order contributions to the fixed-order 
coupling-constant perturbative expansion, on the one hand, and an increasingly 
sophisticated determination of the Parton Distribution Functions (PDFs) of the 
proton~\cite{Gao:2017yyd,Ethier:2020way}, on the other hand.

@@

The consistent inclusion of QCD and EW corrections in precision computations for
LHC processes entails
the solution of two separate problems. First, a problem of efficiency:
fast-interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are essential
whenever the evaluation of the hadronic cross section needs to 
be performed a large number of times, as is the case in the evaluation of
scale variations or of PDF fits. While two default formats exist for
these grids, \textsc{APPLgrid}~\cite{Carli:2010rw} and
\textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}, none of them supports the inclusion of EW
corrections nor the interface to a Monte Carlo generator accurate to 
NLO QCD+EW\@. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such (e.g.\ the $t$-channel photon-induced component in
neutral-current Drell--Yan, which is not a separate processes beyond leading
order) or of just a part of the EW effects (e.g.\ multiple-photon
radiation from light particles in the final state of neutral- or charged-current
Drell--Yan, especially with electrons). Be that as it may, if EW effects are
systematically included in theoretical predictions, they should not be subtracted
from experimental results, otherwise they will be double counted.

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows any user to generate
fast-interpolation grids, accurate to up to any fixed order in the QCD and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation, and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed-orders in the strong coupling only) that can be generated 
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{mg5\_aMC}
henceforth), with which it has been developed and tested. In this respect,
\textsc{PineAPPL} supersedes
\textsc{APPLgrid}+\textsc{aMCfast}~\cite{Bertone:2014zva}.
\textsc{PineAPPL} can also be easily interfaced to any Monte Carlo generator,
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it complements
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.

>>>>

With the recent completion of Run II, the Large Hadron Collider (LHC) has 
accumulated data from an integrated luminosity of approximately 
\SI{150}{\per\femto\barn}~\cite{Mangano:2020icy}. This represents only a small fraction of
the anticipated \SI{3000}{\per\femto\barn} that will eventually be recorded in the
forthcoming twenty years of LHC operation. Nevertheless the statistical 
uncertainty of the data has already shrunk to unprecedentedly small values,
typically \SI{1}{\percent} or less, a fact that will allow for precision tests of the
Standard Model (SM) and for indirect searches of New Physics only if 
theoretical predictions become comparatively precise. This entails the 
computation of additional higher-order contributions to the fixed-order 
perturbative expansion, on the one hand, and an increasingly
sophisticated determination of the Parton Distribution Functions (PDFs) of the 
proton~\cite{Gao:2017yyd,Ethier:2020way}, on the other hand.

@@

The consistent inclusion of QCD and EW corrections in precision computations for
LHC processes entails
the solution of two separate problems. First, a problem of efficiency:
fast-interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are essential
whenever the evaluation of the hadronic cross section needs to 
be performed a large number of times, as is the case in the evaluation of
scale variations or of PDF fits. While two default formats exist for
these grids, \textsc{APPLgrid}~\cite{Carli:2010rw} and
\textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}, none of them supports the inclusion of EW
corrections nor the interface to a Monte Carlo generator accurate to 
NLO QCD+EW\@. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such (e.g.\ the $t$-channel photon-induced component in
neutral-current Drell--Yan, which is not a separate process beyond leading
order) or of just a part of the EW effects (e.g.\ multiple-photon
radiation from light particles in the final state of neutral- or charged-current
Drell--Yan, especially with electrons). Be that as it may, if EW effects are
systematically included in theoretical predictions, they should not be subtracted
from experimental results, otherwise they will be double counted.

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows any user to generate
fast-interpolation grids, accurate to up to any fixed order in the QCD and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation, and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed-orders in the strong coupling only) that can be generated 
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{mg5\_aMC}
henceforth), with which it has been developed and tested. In this respect,
\textsc{PineAPPL} supersedes
\textsc{APPLgrid}+\textsc{aMCfast}~\cite{Bertone:2014zva}.
However, \textsc{PineAPPL} can also be easily interfaced to any Monte Carlo generator,
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it would complement
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.
